Parameter sensitivity and identifiability for a biogeochemical model of hypoxia in the northern {G}ulf of {M}exico\tnoteref{mytitlenote}

### 2016-12-21 ident_manu.Rnw
bib journals now abbreviated

@@

% get online bib file
<<echo = FALSE, cache = FALSE>>=
refs <- httr::GET('https://raw.githubusercontent.com/fawda123/refs/master/refs.bib')
refs <- rawToChar(refs$content)
writeLines(refs, con = file('refs.bib'))
@

@@

\begin{abstract}
\noindent Bio-geo-chemical models are useful tools in environmental sciences that can guide management and policy-making. Consequently, significant time and resources are spent developing these models in system-specific contexts. The optimization of model parameters to maximize precision, including transferability of these models to different systems, are fundamental concerns in the development and application of these tools. This study provides a context for understanding quantitative limitations of coupled hydrodynamic-ecological models by evaluating parameter sensitivity and identifability of a \ac{zerod} unit of a larger spatio-temporal model of hypoxia on the \acl{lcs} of Gulf of Mexico.  The analysis provides a contrast of numeric and ecological certainty in parameter subsets using a systematic framework to infer larger trends in dissolved oxygen dynamics over time, having implications for understanding factors that contribute to environmental conditions that are detrimental to aquatic resources.  In particular, we focus on issues of parameter identifiability using local sensitivity analyses to provide quantitative descriptions of numerical constraints on model precision.  The sensitivity of state variables differed considerably with parameter changes, although most variables were responsive to changes in parameters that influenced planktonic growth rates.  Variation in sensitivity had a direct correspendence with identifiability, such that only small subsets of the complete parameter set were characterized as having unique effects on the model output. As a result, we provide a set of parameter selection heuristics that can be used to identify parameters for model calibration that depend on relative sensitivity and ecological categories within the biogeochemical equations. Although these concerns have been expressed in the literature, they are rarely explicitly addressed or included in evaluations of water quality models.  In addition to immediate implications for regional models, we provide a framework for describing the effects of parameter uncertainty and identifiability that can be applied to similar models to better inform environmental management.
\end{abstract}

@@

\bibliography{refs}
\clearpage

>>>>

\journal{Ecological Modelling}

@@

% get online bib file
<<echo = FALSE, cache = FALSE>>=
refs <- httr::GET('https://raw.githubusercontent.com/fawda123/refs/master/refs.bib')
refs <- rawToChar(refs$content)
writeLines(refs, con = file('refs.bib'))

# library(Jabbrev)
# bbl_ext('ident_manu.bbl', 'refs.bib', 'ident_manu.bib')
# bib_shr('ident_manu.bib', 'refsshr.bib')
@

@@

\begin{abstract}
\noindent The optimization of parameters and transferability to different systems are fundamental concerns in the application of biogeochemical models. This study provides a context for understanding quantitative limitations of coupled hydrodynamic-ecological models by evaluating parameter sensitivity and identifability of a \ac{zerod} unit of a larger spatio-temporal model of hypoxia on the \acl{lcs} of Gulf of Mexico.  The analysis provides a contrast of numeric and ecological certainty in parameter subsets using a systematic framework to infer larger trends in dissolved oxygen dynamics over time, having implications for understanding factors that contribute to environmental conditions that are detrimental to aquatic resources.  In particular, we focus on issues of parameter identifiability using local sensitivity analyses to provide quantitative descriptions of numerical constraints on model precision.  The sensitivity of state variables differed considerably with parameter changes, although most variables were responsive to changes in parameters that influenced planktonic growth rates.  Variation in sensitivity had a direct correspendence with identifiability, such that only small subsets of the complete parameter set were characterized as having unique effects on the model output. As a result, we provide a set of parameter selection heuristics that can be used to identify parameters for model calibration that depend on relative sensitivity and ecological categories within the biogeochemical equations. Although these concerns have been expressed in the literature, they are rarely explicitly addressed or included in evaluations of water quality models.  In addition to immediate implications for regional models, we provide a framework for describing the effects of parameter uncertainty and identifiability that can be applied to similar models to better inform environmental management.
\end{abstract}

@@

\bibliography{refsshr}
\clearpage

### 2016-12-21 ident_manu.tex
bib journals now abbreviated

% get the version based on commit date

@@

\title{Parameter sensitivity and identifiability for a biogeochemical model of hypoxia in the northern {G}ulf of {M}exico\tnoteref{mytitlenote}}
\tnotetext[mytitlenote]{Version: Thu Dec 15 15:47:19 2016 -0600, \href{https://github.com/fawda123/identifiability/commit/f47752c773b8b0d1c90abe57d41f57038201ea42}{f47752c773b8b0d1c90abe57d41f57038201ea42}}

>>>>

% get the version from git log

@@

\title{Parameter sensitivity and identifiability for a biogeochemical model of hypoxia in the northern {G}ulf of {M}exico\tnoteref{mytitlenote}}
\tnotetext[mytitlenote]{Version: Fri Dec 16 10:08:47 2016 -0600, \href{https://github.com/fawda123/identifiability/commit/5e43ad8aba3a34ec8163993b8c3ffa785e1f0fcd}{5e43ad8aba3a34ec8163993b8c3ffa785e1f0fcd}}
